---
title: "descriptive_final"
author: "Abhishek U Verma"
date: "May 3, 2017"
output: html_document
---

##I - Loading the data

```{r}
setwd('D:/College Work/Semester_1/Data Warehousing, Mining & Vis.(IS-665)/Notes/Project')
my.dataset=read.csv('CarsDataset.csv')
dim(my.dataset)
test = my.dataset
```


##II - Exploratory Data Analysis


```{r}
dim(test)
names(test)
head(test)
tail(test)
```

```{r}
str(test)
```


##III - Looking for Missing Values

```{r}
sum(is.na(test))

```


##IV - Descriptive Statistics for importants variables

Engine displacement: It is a size of engine in terms of cubic centimeter.  

```{r}
typeof(test$engine_displacement) #is.numeric(yps$Age)
summary(test$engine_displacement) #Youngest participants have 15yrs and oldest 30yrs. Typical participant has 20 years of age.
table(test$engine_displacement) #Distribution of participants by Age
table(test$engine_displacement,test$engine_power) #Distribution of participants by age and by gender

which(is.na(test$engine_displacement)) #identify rows(observations) with NA for Age

sum(is.na(test$engine_displacement))

sd(test$engine_displacement, na.rm = TRUE) #To remove missing values (NA'S) and avoid gettin na as sd

IQR(test$engine_displacement, na.rm = TRUE)

displacement_mean = mean(test$engine_displacement)
displacement_median = median(test$engine_displacement)

table(test$engine_displacement,test$price_eur) #price level by engine displacement.

skewness_displacement=3*(displacement_mean-displacement_median)/sd(test$engine_displacement) # Data is positively skewed
skewness_displacement
```


```{r}
typeof(test$engine_power) #is.numeric(yps$Age)
summary(test$engine_power) #Youngest participants have 15yrs and oldest 30yrs. Typical participant has 20 years of age.
table(test$engine_power) #Distribution of participants by Age

which(is.na(test$engine_power)) #identify rows(observations) with NA for Age

sum(is.na(test$engine_power))

sd(test$engine_power, na.rm = TRUE) #To remove missing values (NA'S) and avoid gettin na as sd

IQR(test$engine_power, na.rm = TRUE)

power_mean = mean(test$engine_power)
power_median = median(test$engine_power)

test$engine_power[which(is.na(test$engine_power))] <- power_mean #replace na of Age variable by the mean

skewness_power=3*(power_mean-power_median)/sd(test$engine_power)# Data is positively skewed
skewness_power
```

##V- Interpretation: 

Ref-1 : Here we can see that the data for the displacement variable is significantly skewed to the right. Looking at the histogram we can infer that the majority of the cars range from 1000 to 3000 cubic centimeter.

Ref-2 : Here we can see that the data for the engine power variable is significantly skewed to the right. This histogram shows that teh majority of the cars range from 0-200 horsepower.

```{r}
#Ref - 1
hist(test$engine_displacement, main = 'engine displacement Statistics', col = 'salmon', xlab = 'displacement', ylab = 'Number of cars') 

#Ref - 2

hist(test$engine_power, main = 'engine power Statistics', col = 'salmon', xlab = 'power', ylab = 'Number of cars') 

```



##VI- Handling of missing values for Gender variable

As there are no missing values thats why this step is skipped in descriptive statistics

##VII- Describing categorical values

```{r}

summary(test$transmission)
summary(test$fuel_type)

man_power = test[test$transmission=='man',]$engine_power
auto_power = test[test$transmission=='auto',]$engine_power

gas_power = test[test$fuel_type=='gasoline',]$engine_power
diesel_power = test[test$fuel_type=='diesel',]$engine_power

mean(man_power, na.rm=T)
mean(auto_power, na.rm = T)

mean(gas_power, na.rm = T)
mean(diesel_power, na.rm = T)

```


```{r}
barplot(table(test$transmission))
barplot(table(test$fuel_type))

```


